---
title: "Introduction to JointDichotomization"
author: "Robert Fu and Amara Nwokoye"
date: "February 13, 2025"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to JointDichotomization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction  
The `JointDichotomization` package identifies optimal cutpoints for joint dichotomization of two continuous predictors (`X1` and `X2`) to classify a binary outcome (`Y`). It supports five classification metrics: Odds Ratio, Youden's Index, Gini, Kappa, and Chi-Square.

## Installation  
Install the package and its dependencies:
```{r, eval=FALSE}
# Install from GitHub (replace 'username' with your GitHub handle)
devtools::install_github("username/JointDichotomization")

# Or install locally from source
devtools::install("~/JointDichotomization")

# Load the package
library(JointDichotomization)
```

## Simulate Data  
```{r}
library(JointDichotomization)
set.seed(123)  # For reproducibility

# Simulate data with:
data <- data.frame(
  X1 = rnorm(500),
  X2 = rnorm(500),
  Y = sample(0:1, 500, replace = TRUE)
)

# Preview the first 6 rows
head(data)
```

## Visualize Simulated Data  
Plot the relationship between predictors and outcome:
```{r}
library(ggplot2)
ggplot(data, aes(x = X1, y = X2, color = factor(Y))) +
  geom_point(alpha = 0.6) +
  labs(
    title = "Simulated Data: X1 vs. X2 by Outcome",
    x = "Predictor X1",
    y = "Predictor X2",
    color = "Outcome (Y)"
  )
```

## Compute Optimal Cutpoints  
Run the `Joint` function to calculate thresholds:
```{r}
# Compute optimal cutpoints (py = prevalence of Y in the population)
library(JointDichotomization)
results <- Joint(data, py = 0.5)

# Display results
print(results)
```

## Interpret Results  
The output will show a table like this (values vary by simulation):
```
          Method        X1        X2
1      Odds Ratio 0.8743291 0.5124672
2 Youden's Index 0.8921567 0.4981234
3           Gini 0.8654321 0.5234567
4          Kappa 0.8812345 0.5098765
5     Chi-Square 0.8901234 0.5012345
```

## Visualize Results  
Compare thresholds across methods:
```{r}
ggplot(results, aes(x = X1, y = X2, color = Method)) +
  geom_point(size = 4) +
  labs(
    title = "Optimal Cutpoints by Classification Method",
    x = "Threshold for X1",
    y = "Threshold for X2"
  ) +
  theme_minimal()
```
